Commutative character sheaves and geometric types for supercuspidal representations

### 2017-05-11 Geometric types.tex
Removing comments, overful hbox fixes

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups can be built from ``data'' introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu data can be constructed directly from the datum, it is convenient to consider an intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
%
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations 
%[appearing in Yu type data]
 are on groups of the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over [a henselian discrete valuation ring with finite residue field $\kappa$] $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} %\cite{Yu:models}
\end{quotation}
%
In this paper we follow the suggestion above by showing that Yu type data are geometrizable, in the following sense.
A Yu type datum determines a sequence of representations $\orho_i$ of compact $p$-adic groups $\oK^i$, for $i=0, \ldots, d$, such that $(\oK^d, \rho_d)$ is a type for a supercuspidal representation of a $p$-adic group.
Let $R$ be the ring of integers of a local field with finite residue field $k$.
The main result of \cite{Yu:models} shows how to find, for each $i=0, \ldots, d$, a smooth group scheme $\underline{G}^i$ over the ring $R$ with $\underline{G}^i(R)=\,^\circ K^i $.
The geometrization of the Yu type datum uses Lusztig's theory of character sheaves on reductive groups over finite fields, so it is necessary to assume that the geometric component group of the reductive quotient of the special fibre of the group scheme $\underline{G}^0$ is cyclic, in order to bring his work to bear. Under this assumption we show how each representation $\orho_i$ can be replaced by a pair $(\underline{G}^i, \cs{F}^i)$, where $\cs{F}^i$ is a {\it rational virtual} sheaf complex on the Greenberg transform $G^i$ of $\underline{G}^i$, by which we mean $\cs{F}^i$ is an element in the group obtained by tensoring the Grothendieck group of such sheaves with $\QQ$. 
Writing $\trFrob{\cs{F}^i}$ for the function on $G^i(k) = \underline{G}^i(R) = \,^\circ K^i$ obtained by evaluating the trace of the action of Frobenius on the rational virtual complex $\cs{F}^i$, we show in Theorem~\ref{thm:geotypes} that 
\begin{equation}\label{eqn:intro1}
\trFrob{\cs{F}^i} = \trace(\,^\circ\rho_i).
\end{equation}
By this theorem, then, we obtain geometric avatars for each type in a Yu datum:
\[
\begin{tikzcd}
(\oK^i,\orho_i) \arrow[bend left, dashed]{rrr}{\text{geometrization}} &&& \arrow[bend left, dashed]{lll}{\text{trace of Frob}}
(\underline{G}^i, \cs{F}^i).
\end{tikzcd}
\]
%Theorem~\ref{thm:geotypes} shows how to construct the rational virtual Weil sheaf $\cs{F}^i$ from the Yu type datum, so that \eqref{eqn:intro1} holds.
We refer to the pair $(\underline{G}^d, \cs{F}^d)$ as a \emph{geometric type}.

@@

The function-sheaf dictionary over a finite field $k$ \cite{deligne:SGA4.5}*{Sommes trig.}
provides a way of encoding functions on the $k$-rational points of an algebraic group $G$
as $\ell$-adic local systems on $G$.  More specifically, if $G$ is a connected, commutative, algebraic group
then there is a certain category $\CS(G)$ of rank-one local systems on $G$ and an
explicit isomorphism between isomorphism classes
of objects in $\CS(G)$ and $G(k)^* \ceq \Hom(G(k), \EEx)$; 
%Objects in $\CS(G)$ are referred to as \emph{character sheaves}.
the isomorphism is given by mapping $\cs{L}$ to the function
$\TrFrob{G} : g \mapsto \Tr(\Frob{} \vert \cs{L}_g)$.

@@

Next, we broaden our scope further to encompass smooth group schemes $G$ over $\Fq$ that are not necessarily commutative.
We assume $G$ is smooth, but not that it is connected, reductive or commutative. 
The category $\CS(G)$ has a straightforward generalization to this case, but again
there are more character sheaves than there are characters, as pointed out by Kamgarpour \cite{kamgarpour:09a}*{(1.1)}.
We then define category $\CCS(G)$ for such $G$ and a forgetful functor to $\CS(G)$ so that $\TrFrob{G} : \CCSiso{G} \to G\ab(k)^*$
is an isomorphism.  
Since $G\ab(k)^*$ surjects onto $G(k)^*$, it follows that for each character $\chi \in G(k)^*$ there is a commutative
character sheaf $\cs{L}$ on $G$ with $\TrFrob{G}(\cs{L}) = \chi$. Moreover, we find that pullback along the quotient $q : G \to G\ab$
defines an equivalence of categories $\CCS(G\ab) \to \CCS(G)$.  
The functor $\CCS(G) \to \CS(G)$ is not
essentially surjective, missing the kinds of linear character sheaves highlighted by Kamgarpour.
%

@@

We now summarize the sections of the paper in more detail.
%
In Section \ref{sec:defs}, we recall the category $\CS(G)$ from \cite{cunningham-roe:13a} and note that it still makes sense when $G$ is not commutative.  
%
We focus on the case of commutative $G$ in Section \ref{sec:comcom},
giving the definition of a commutative character sheaf and proving our first main theorem, that
$\TrFrob{G} : \CSiso{G} \to G(k)^*$ induces an isomorphism on $\CCSiso{G}$.
%
Passing to the case that $G$ is non-commutative, we give the definition of and main results about commutative character sheaves in Section \ref{sec:noncom}.  
We note that we should only consider character sheaves that arise via pullback from $G\ab$ in order to eliminate those that have nontrivial restriction to the derived subgroup.  
This observation underlies the definition of commutative character sheaves for non-commutative $G$.  
We state our second main result, Theorem~\ref{thm:geo}, that pullback along the abelianization map defines an equivalence of categories $\CCS(G) \to \CCS(G\ab)$.
In Section \ref{ssec:obmor}, we use Galois cohomology to describe the relationship between $G(k)^*$ and $G\ab(k)^*$.  
We also compute the automorphism groups in $\CCS(G)$.
%Then in Section \ref{ssec:proofs}, we give proofs of the results in Section \ref{sec:noncom}, which require a development of equivariant linear character sheaves. \todo{Fix this.}
% 
In Section~\ref{sec:types} we use Theorem~\ref{thm:geo} to geometrize types for supercuspidal representations of $p$-adic groups, in a sense made precise in Theorem~\ref{thm:geotypes}. 
% is given in Section~\ref{ssec:geotypes}. 
 %Specifically, we show that if $C$ is a compact $p$-adic group and $\rho$ is a representation of $C$ and $(C,\rho)$ is a type for a tamely ramified $p$-adic group then there is a  group scheme $G$ over $\Fq$ such that $G(\Fq) = C$ and a rational virtual Weil sheaf $\mathcal{A}$ on $G$ such that $\trFrob{\cs{A}} = \trace(\rho)$.
As preparation for the proof, we review some facts about the Heisenberg-Weil representation and its geometrization, in Section~\ref{ssec:Jacobi}.
Then, in Section~\ref{ssec:review}, we review Yu's theory of types and his study of smooth integral models.  
These elements are pulled together in Section~\ref{ssec:geotypes}, where the proof Theorem~\ref{thm:geotypes} is given.

@@

In order to get character sheaves that correspond to characters in $G(k)^*$, we must discard those
character sheaves whose restriction to the derived subgroup is nontrivial.
%
Recall from Section~\ref{sec:defs} that we refer to objects in category $\CS(G)$, defined as
in \cite{cunningham-roe:13a}*{Def. 1.1}, as linear character sheaves when $G$ is smooth but
not necessarily commutative.  
We define the following category to track the trivialization on the derived subgroup;
commutative character sheaves will then be defined as a subcategory.

@@

The reason for tracking $\beta$ is that it determines an isomorphism $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ as follows,
which will replace the $\xi$ of Definition \ref{def:CCScom}.
Let $i : G \to G$ be inversion and $c : G\times G\to G\der$ be the commutator map, defined by $c(x,y)= xyx^{-1}y^{-1}$.
Both are smooth morphisms of $\Fq$-schemes.
Set $m' = i \circ m \circ \theta$ and let $j\der : G\der\to G$ be inclusion; then $j\der\circ c = m \circ (m \times m')$.
% \todo{$j$ is not defined, and the needed definition conflicts with the definition of $i$ and $j$ given later.}
Then $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ by the diagram of isomorphisms below.
\begin{equation}
\begin{tikzcd}
\arrow[equal]{d} c^* (\cs{L}\vert_{G\der}) \arrow{r}{c^*(\beta)} 
	&  c^*((\EE)_{G\der}) \arrow[equal]{d} \\
\arrow{d}[swap]{j\der\circ c = m \circ (m \times m')} c^* j\der^* \cs{L} 
	&   (\EE)_{G\times G} \\
(m \times m')^* m^* \cs{L} \arrow{d}[swap]{(m \times m')^*(\mu)} 
	&  m^*\cs{L} \otimes \theta^* m^* i^* \cs{L}  \arrow[dashed]{u}[swap]{\gamma'} \\
(m \times m')^* (\cs{L} \boxtimes \cs{L}) \arrow[equal]{r} 
	& m^*\cs{L} \otimes (m')^*\cs{L} \arrow{u}[swap]{m' = i\circ m\circ \theta} 
\end{tikzcd}
\end{equation}
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations; so, for example, the middle left isomorphism comes from $(m\times m')^* m^* \iso c^* j\der^*$ since $j\der\circ c = m \circ (m \times m')$.
Using the monoidal structure of the category of Weil local systems on $G\times G$, the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ defines an isomorphism
\[
m^*\cs{L} \to (\theta^* m^* i^*\cs{L})^\vee.
\]
Applying the canonical isomorphisms $(\theta^* m^* i^*\cs{L})^\vee \iso \theta^* m^* i^* (\cs{L}^\vee)$ and $i^*(\cs{L}^\vee) \cong \cs{L}$, this map provides the promised isomorphism
\[
\begin{tikzcd}
\gamma : m^*\cs{L} \arrow{r} & \theta^* m^* \cs{L}.
\end{tikzcd}
\]

@@

Let $\Loc(G)$ and $\Loc(G\ab)$ be the categories of Weil local systems on $G$ and $G\ab$, respectively.
%
Let $\Loc\der(G)$ be the category of $G\der$-equivariant Weil local systems on $G$, whose definition we now recall.
Let $n : G\der\times G\to G$ be the restriction of $m : G\times G\to G$ to $G\der\times G$,
let $p : G\der\times G \to G$ be projection to the second component, and let $s: G \to G\der\times G$ be given by $s(g) = (1,g)$.
Then the quotient $q : G \to G\ab$ is a regular epimorphism of smooth group schemes with kernel pair $(n,p)$.
\[
\begin{tikzcd}
G\der\times G
 \arrow[shift left=1]{r}{n}
  \arrow[shift right=1,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G\ab
\end{tikzcd}
\]
Consider the morphisms
\[
\begin{tikzcd}
G\der \times G\der \times G 
\arrow[shift left=2]{r}{b_1, b_2, b_3} 
\arrow{r}{}
\arrow[shift right=2]{r}{} 
& G\der \times G 
\arrow[shift left=1]{r}{n}
\arrow[shift right=1]{r}[swap]{p}
 & G 
\end{tikzcd}
\]
defined by 
\begin{align*}
b_1(h_1,h_2,g) &= (h_1h_2,g) \\
b_2(h_1,h_2,g) &= (h_1,h_2g) \\
b_3(h_1,h_2,g) &= (h_2,g).
\end{align*}
Note that
\begin{align*}
n\circ b_1 &= n\circ b_2 \\
n\circ b_3 &= p\circ b_2 \numberthis \label{eqn:bap}\\
p\circ b_1 &= p\circ b_3.
\end{align*}
A $G\der$-equivariant Weil local system on $G$ is a Weil local system $\cs{L}$ on $G$ together with an isomorphism  
\[
\nu : n^*\cs{L} \to p^*\cs{L}
\] 
of Weil local systems on $G\der\times G$ such that 
\begin{equation}\label{E1}
s^*(\nu) = \id_{\cs{L}}
\end{equation}
and such that the following diagram of isomorphisms of local systems on $G\der\times G\der\times G$ commutes. 
\begin{equation}\label{E2}
\begin{tikzcd}
\ &  \arrow{dl}{n\circ b_1 = n\circ b_2}  b_2^*  n^*\, \mathcal{L} \arrow{rr}{b_2^*(\nu)} && b_2^* p^*\, \mathcal{L} \arrow{dr}[swap]{p\circ b_2 = n\circ b_3} & \\
b_1^* n^*\, \mathcal{L} \arrow{dr}{b_1^*(\nu)} &&&&   \arrow{dl}[swap]{b_3^*(\nu)} b_3^* n^*\, \mathcal{L} \\
& b_1^* p^*\, \mathcal{L}  && \arrow{ll}[swap]{p\circ b_3 = p\circ b_1} b_3^* p^*\, \mathcal{L} & 
\end{tikzcd}
\end{equation}
%In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from equations \eqref{eqn:bap}; so, for example, the top left isomorphism comes from $b_1^* n^* \iso b_2^* n^*$ since $n\circ b_1 = n\circ b_2$.
%On stalks, this condition is ...
%\[
%\begin{tikzcd}
%\ &  \cs{L}_{h_1\cdot(h_2\cdot g)} \arrow{r}{\nu_{h_1,h_2\cdot g}} & \cs{L}_{h_2\cdot g} & \\
%\cs{L}_{(h_1 h_2) \cdot g} \arrow{dr}[swap]{\nu_{(h_1h_2),g}} \arrow[equal]{ur} &&& \arrow[equal]{ul}  \arrow{dl}{\nu_{h_2,g}} \cs{L}_{h_2\cdot g} \\
%& \cs{L}_g \arrow[equal]{r} & \cs{L}_g & 
%\end{tikzcd}
%\]
Morphisms of $H$-equivariant Weil local systems $(\cs{L},\nu)\to (\cs{L}',\nu')$ are morphisms of Weil local systems $\alpha: \cs{L}\to \cs{L}'$ for which the diagram
\begin{equation}\label{E3}
\begin{tikzcd}
\arrow{d}[swap]{\nu} n^*\cs{L} \arrow{r}{n^*(\alpha)} & n^*\cs{L}' \arrow{d}{\nu'} \\
p^*\cs{L} \arrow{r}{p^*(\alpha)} & p^*\cs{L}'
\end{tikzcd}
\end{equation}
commutes.
This defines $\Loc\der(G)$, the category of $G\der$-equivariant Weil local systems on $G$.
The reader will recognize this notion as the Weil local system version of equivariant sheaves for the action $n$ of $G\der$ on $G$, as can be found, for example, in \cite{bernstein-lunts:equivariant}*{0.2}.

@@

With reference to Section~\ref{ssec:equivariant1}, we define a $G\der$-equivariant linear character sheaf on $G$ to be a triple $(\cs{L},\mu, \nu)$, where $(\cs{L},\mu)$ is a linear character sheaf and $(\cs{L},\nu)$ is an $G\der$-equivariant Weil local system.  
A morphism of $G\der$-equivariant linear character sheaves $(\cs{L},\mu,\nu) \to (\cs{L}',\mu',\nu')$ is a morphism of $G\der$-equivariant Weil local systems $\alpha : \cs{L}\to \cs{L}'$ which is also a morphism of linear character sheaves.
%so
%\begin{equation}\label{CS4}
%\begin{tikzcd}
%\arrow{d}[swap]{\mu} m^*\cs{L} \arrow{r}{m^*(\alpha)} & m^*\cs{L}' \arrow{d}{\mu'} \\
%\cs{L}\boxtimes \cs{L} \arrow{r}{\alpha\boxtimes\alpha} & \cs{L}'\boxtimes\cs{L}' 
%\end{tikzcd}
%\end{equation}
%commutes; see also \cite{cunningham-roe:13a}, CS.4].
%
Let $\CS\der(G)$ be the category of $G\der$-equivariant linear character sheaves on $G$.

%Is $\CS_H(G)$ a rigid monoidal category?

%If $(\cs{L},\mu,\nu)$ and $(\cs{L},\mu,\nu')$ in $\CS_H(G)$, then $\nu = \nu'$?

@@

\begin{lemma}\label{lem:HH}
If $(\cs{L},\mu,\nu)$ is a $G\der$-equivariant linear character sheaf on $G$ then $\mu : m^*\cs{L} \to \cs{L}\boxtimes \cs{L}$ 
%and $\vartheta : \cs{L}\boxtimes \cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$ 
is a morphism of $G^2\der$-equivariant Weil local systems on $G^2$. 
\end{lemma}

>>>>

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups can be built from ``data'' introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu data can be constructed directly from the datum, it is convenient to consider an intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations 
 are on groups of the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over [a henselian discrete valuation ring with finite residue field $\kappa$] $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.}
\end{quotation}
In this paper we follow the suggestion above by showing that Yu type data are geometrizable, in the following sense.
A Yu type datum determines a sequence of representations $\orho_i$ of compact $p$-adic groups $\oK^i$, for $i=0, \ldots, d$, such that $(\oK^d, \rho_d)$ is a type for a supercuspidal representation of a $p$-adic group.
Let $R$ be the ring of integers of a local field with finite residue field $k$.
The main result of \cite{Yu:models} shows how to find, for each $i=0, \ldots, d$, a smooth group scheme $\underline{G}^i$ over the ring $R$ with $\underline{G}^i(R)=\,^\circ K^i $.
The geometrization of the Yu type datum uses Lusztig's theory of character sheaves on reductive groups over finite fields, so it is necessary to assume that the geometric component group of the reductive quotient of the special fibre of the group scheme $\underline{G}^0$ is cyclic, in order to bring his work to bear. Under this assumption we show how each representation $\orho_i$ can be replaced by a pair $(\underline{G}^i, \cs{F}^i)$, where $\cs{F}^i$ is a {\it rational virtual} sheaf complex on the Greenberg transform $G^i$ of $\underline{G}^i$, by which we mean $\cs{F}^i$ is an element in the group obtained by tensoring the Grothendieck group of such sheaves with $\QQ$. 
Writing $\trFrob{\cs{F}^i}$ for the function on $G^i(k) = \underline{G}^i(R) = \,^\circ K^i$ obtained by evaluating the trace of the action of Frobenius on the rational virtual complex $\cs{F}^i$, we show in Theorem~\ref{thm:geotypes} that 
\begin{equation}\label{eqn:intro1}
\trFrob{\cs{F}^i} = \trace(\,^\circ\rho_i).
\end{equation}
By this theorem, then, we obtain geometric avatars for each type in a Yu datum:
\[
\begin{tikzcd}
(\oK^i,\orho_i) \arrow[bend left, dashed]{rrr}{\text{geometrization}} &&& \arrow[bend left, dashed]{lll}{\text{trace of Frob}}
(\underline{G}^i, \cs{F}^i).
\end{tikzcd}
\]
We refer to the pair $(\underline{G}^d, \cs{F}^d)$ as a \emph{geometric type}.

@@

The function-sheaf dictionary over a finite field $k$ \cite{deligne:SGA4.5}*{Sommes trig.}
provides a way of encoding functions on the $k$-rational points of an algebraic group $G$
as $\ell$-adic local systems on $G$.  More specifically, if $G$ is a connected, commutative, algebraic group
then there is a certain category $\CS(G)$ of rank-one local systems on $G$ and an
explicit isomorphism between isomorphism classes
of objects in $\CS(G)$ and $G(k)^* \ceq \Hom(G(k), \EEx)$; 
the isomorphism is given by mapping $\cs{L}$ to the function
$\TrFrob{G} : g \mapsto \Tr(\Frob{} \vert \cs{L}_g)$.

@@

Next, we broaden our scope further to encompass smooth group schemes $G$ over $\Fq$ that are not necessarily commutative.
We assume $G$ is smooth, but not that it is connected, reductive or commutative. 
The category $\CS(G)$ has a straightforward generalization to this case, but again
there are more character sheaves than there are characters, as pointed out by Kamgarpour \cite{kamgarpour:09a}*{(1.1)}.
We then define category $\CCS(G)$ for such $G$ and a forgetful functor to $\CS(G)$ so that $\TrFrob{G} : \CCSiso{G} \to G\ab(k)^*$
is an isomorphism.  
Since $G\ab(k)^*$ surjects onto $G(k)^*$, it follows that for each character $\chi \in G(k)^*$ there is a commutative
character sheaf $\cs{L}$ on $G$ with $\TrFrob{G}(\cs{L}) = \chi$. Moreover, we find that pullback along the quotient $q : G \to G\ab$
defines an equivalence of categories $\CCS(G\ab) \to \CCS(G)$.  
The functor $\CCS(G) \to \CS(G)$ is not
essentially surjective, missing the kinds of linear character sheaves highlighted by Kamgarpour.

@@

We now summarize the sections of the paper in more detail.
In Section \ref{sec:defs}, we recall the category $\CS(G)$ from \cite{cunningham-roe:13a} and note that it still makes sense when $G$ is not commutative.  
We focus on the case of commutative $G$ in Section \ref{sec:comcom},
giving the definition of a commutative character sheaf and proving our first main theorem, that
$\TrFrob{G} : \CSiso{G} \to G(k)^*$ induces an isomorphism on $\CCSiso{G}$.
Passing to the case that $G$ is non-commutative, we give the definition of and main results about commutative character sheaves in Section \ref{sec:noncom}.  
We note that we should only consider character sheaves that arise via pullback from $G\ab$ in order to eliminate those that have nontrivial restriction to the derived subgroup.  
This observation underlies the definition of commutative character sheaves for non-commutative $G$.  
We state our second main result, Theorem~\ref{thm:geo}, that pullback along the abelianization map defines an equivalence of categories $\CCS(G) \to \CCS(G\ab)$.
In Section \ref{ssec:obmor}, we use Galois cohomology to describe the relationship between $G(k)^*$ and $G\ab(k)^*$.  
We also compute the automorphism groups in $\CCS(G)$.
In Section~\ref{sec:types} we use Theorem~\ref{thm:geo} to geometrize types for supercuspidal representations of $p$-adic groups, in a sense made precise in Theorem~\ref{thm:geotypes}.
As preparation for the proof, we review some facts about the Heisenberg-Weil representation and its geometrization, in Section~\ref{ssec:Jacobi}.
Then, in Section~\ref{ssec:review}, we review Yu's theory of types and his study of smooth integral models.  
These elements are pulled together in Section~\ref{ssec:geotypes}, where the proof Theorem~\ref{thm:geotypes} is given.

@@

In order to get character sheaves that correspond to characters in $G(k)^*$, we must discard those
character sheaves whose restriction to the derived subgroup is nontrivial.
Recall from Section~\ref{sec:defs} that we refer to objects in category $\CS(G)$, defined as
in \cite{cunningham-roe:13a}*{Def. 1.1}, as linear character sheaves when $G$ is smooth but
not necessarily commutative.  
We define the following category to track the trivialization on the derived subgroup;
commutative character sheaves will then be defined as a subcategory.

@@

The reason for tracking $\beta$ is that it determines an isomorphism $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ as follows,
which will replace the $\xi$ of Definition \ref{def:CCScom}.
Let $i : G \to G$ be inversion and $c : G\times G\to G\der$ be the commutator map, defined by $c(x,y)= xyx^{-1}y^{-1}$.
Both are smooth morphisms of $\Fq$-schemes.
Set $m' = i \circ m \circ \theta$ and let $j\der : G\der\to G$ be inclusion; then $j\der\circ c = m \circ (m \times m')$.
Then $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ by the diagram of isomorphisms below.
\begin{equation}
\begin{tikzcd}
\arrow[equal]{d} c^* (\cs{L}\vert_{G\der}) \arrow{r}{c^*(\beta)} 
	&  c^*((\EE)_{G\der}) \arrow[equal]{d} \\
\arrow{d}[swap]{j\der\circ c = m \circ (m \times m')} c^* j\der^* \cs{L} 
	&   (\EE)_{G\times G} \\
(m \times m')^* m^* \cs{L} \arrow{d}[swap]{(m \times m')^*(\mu)} 
	&  m^*\cs{L} \otimes \theta^* m^* i^* \cs{L}  \arrow[dashed]{u}[swap]{\gamma'} \\
(m \times m')^* (\cs{L} \boxtimes \cs{L}) \arrow[equal]{r} 
	& m^*\cs{L} \otimes (m')^*\cs{L} \arrow{u}[swap]{m' = i\circ m\circ \theta} 
\end{tikzcd}
\end{equation}
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations; so, for example, the middle left isomorphism comes from $(m\times m')^* m^* \iso c^* j\der^*$ since $j\der\circ c = m \circ (m \times m')$.
Using the monoidal structure of the category of Weil local systems on $G\times G$, the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ defines an isomorphism
\[
m^*\cs{L} \to (\theta^* m^* i^*\cs{L})^\vee.
\]
Applying the canonical isomorphisms $(\theta^* m^* i^*\cs{L})^\vee \iso \theta^* m^* i^* (\cs{L}^\vee)$ and $i^*(\cs{L}^\vee) \cong \cs{L}$, this map provides the promised isomorphism
\[
\begin{tikzcd}
\gamma : m^*\cs{L} \arrow{r} & \theta^* m^* \cs{L}.
\end{tikzcd}
\]

@@

Let $\Loc(G)$ and $\Loc(G\ab)$ be the categories of Weil local systems on $G$ and $G\ab$, respectively.
Let $\Loc\der(G)$ be the category of $G\der$-equivariant Weil local systems on $G$, whose definition we now recall.
Let $n : G\der\times G\to G$ be the restriction of $m : G\times G\to G$ to $G\der\times G$,
let $p : G\der\times G \to G$ be projection to the second component, and let $s: G \to G\der\times G$ be given by $s(g) = (1,g)$.
Then the quotient $q : G \to G\ab$ is a regular epimorphism of smooth group schemes with kernel pair $(n,p)$.
\[
\begin{tikzcd}
G\der\times G
 \arrow[shift left=1]{r}{n}
  \arrow[shift right=1,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G\ab
\end{tikzcd}
\]
Consider the morphisms
\[
\begin{tikzcd}
G\der \times G\der \times G 
\arrow[shift left=2]{r}{b_1, b_2, b_3} 
\arrow{r}{}
\arrow[shift right=2]{r}{} 
& G\der \times G 
\arrow[shift left=1]{r}{n}
\arrow[shift right=1]{r}[swap]{p}
 & G 
\end{tikzcd}
\]
defined by 
\begin{align*}
b_1(h_1,h_2,g) &= (h_1h_2,g) \\
b_2(h_1,h_2,g) &= (h_1,h_2g) \\
b_3(h_1,h_2,g) &= (h_2,g).
\end{align*}
Note that
\begin{align*}
n\circ b_1 &= n\circ b_2 \\
n\circ b_3 &= p\circ b_2 \numberthis \label{eqn:bap}\\
p\circ b_1 &= p\circ b_3.
\end{align*}
A $G\der$-equivariant Weil local system on $G$ is a Weil local system $\cs{L}$ on $G$ together with an isomorphism  
\[
\nu : n^*\cs{L} \to p^*\cs{L}
\] 
of Weil local systems on $G\der\times G$ such that 
\begin{equation}\label{E1}
s^*(\nu) = \id_{\cs{L}}
\end{equation}
and such that the following diagram of isomorphisms of local systems on $G\der\times G\der\times G$ commutes. 
\begin{equation}\label{E2}
\begin{tikzcd}
\ &  \arrow{dl}{n\circ b_1 = n\circ b_2}  b_2^*  n^*\, \mathcal{L} \arrow{rr}{b_2^*(\nu)} && b_2^* p^*\, \mathcal{L} \arrow{dr}[swap]{p\circ b_2 = n\circ b_3} & \\
b_1^* n^*\, \mathcal{L} \arrow{dr}{b_1^*(\nu)} &&&&   \arrow{dl}[swap]{b_3^*(\nu)} b_3^* n^*\, \mathcal{L} \\
& b_1^* p^*\, \mathcal{L}  && \arrow{ll}[swap]{p\circ b_3 = p\circ b_1} b_3^* p^*\, \mathcal{L} & 
\end{tikzcd}
\end{equation}
Morphisms of $H$-equivariant Weil local systems $(\cs{L},\nu)\to (\cs{L}',\nu')$ are morphisms of Weil local systems $\alpha: \cs{L}\to \cs{L}'$ for which the diagram
\begin{equation}\label{E3}
\begin{tikzcd}
\arrow{d}[swap]{\nu} n^*\cs{L} \arrow{r}{n^*(\alpha)} & n^*\cs{L}' \arrow{d}{\nu'} \\
p^*\cs{L} \arrow{r}{p^*(\alpha)} & p^*\cs{L}'
\end{tikzcd}
\end{equation}
commutes.
This defines $\Loc\der(G)$, the category of $G\der$-equivariant Weil local systems on $G$.
The reader will recognize this notion as the Weil local system version of equivariant sheaves for the action $n$ of $G\der$ on $G$, as can be found, for example, in \cite{bernstein-lunts:equivariant}*{0.2}.

@@

With reference to Section~\ref{ssec:equivariant1}, we define a $G\der$-equivariant linear character sheaf on $G$ to be a triple $(\cs{L},\mu, \nu)$, where $(\cs{L},\mu)$ is a linear character sheaf and $(\cs{L},\nu)$ is an $G\der$-equivariant Weil local system.  
A morphism of $G\der$-equivariant linear character sheaves $(\cs{L},\mu,\nu) \to (\cs{L}',\mu',\nu')$ is a morphism of $G\der$-equivariant Weil local systems $\alpha : \cs{L}\to \cs{L}'$ which is also a morphism of linear character sheaves.
Let $\CS\der(G)$ be the category of $G\der$-equivariant linear character sheaves on $G$.

@@

\begin{lemma}\label{lem:HH}
If $(\cs{L},\mu,\nu)$ is a $G\der$-equivariant linear character sheaf on $G$ then $\mu : m^*\cs{L} \to \cs{L}\boxtimes \cs{L}$ 
is a morphism of $G^2\der$-equivariant Weil local systems on $G^2$. 
\end{lemma}

\begin{proof} 
Define
\begin{align*}
d : G\der\times G\der\times G\times G &\to G\der\times G\times G\der\times G \\
(h_1,h_2,g_1,g_2) &\mapsto (h_1, g_1, h_2, g_2)\\
n_2 : G\der\times G\times G\der\times G &\to G\times G \\
(h_1,g_1,h_2,g_2) &\mapsto (h_1g_1, h_2g_2) \\
p_2 : G\der\times G\times G\der\times G &\to G\times G \\
(h_1,g_1,h_2,g_2) &\mapsto ( g_1,g_2).
\end{align*}
The following diagram defines the isomorphisms needed to see that both $m^*\cs{L}$ and $\cs{L}\boxtimes\cs{L}$ are $G\der^2$-equivariant Weil local systems.
\[
\begin{tikzcd}[column sep=40]
\arrow{d}[swap]{n_2^*(\mu)} n_2^* (m^*\cs{L}) \arrow[dashed]{r} 
	& \arrow{d}{p_2^*(\mu)} p_2^*(m^*\cs{L})\\
\arrow{d}[swap]{n_2= n^2\circ d} n_2^*(\cs{L}\boxtimes\cs{L}) \arrow[dashed]{r} 
	& \arrow{d}{p_2= p^2\circ d} p_2^*(\cs{L}\boxtimes\cs{L}) \\
d^* (n^*\cs{L}\boxtimes n^*\cs{L}) \arrow{r}{d^*(\nu\boxtimes \nu)}& d^*(p^*\cs{L}\boxtimes p^*\cs{L}) 
\end{tikzcd}
\]
The dashed arrows both satisfy \eqref{E1} and \eqref{E2} as they apply here.
This diagram also shows that $\mu : m^*\cs{L} \to \cs{L}\boxtimes \cs{L}$ is a morphism of $G\der^2$-equivariant local systems, since it satisfies \eqref{E3} as it applies here.
\end{proof}
